Palmer amaranth (\emph{Amaranthus palmeri}) adaptation to agroecossystems

### 2021-07-26 Frontiers/Frontiers.Rmd
as

---
title: "Adaptation of Palmer amaranth (*Amaranthus palmeri*) to agroecossystems"
runtitle: "Short Title"
documentclass: frontiersSCNS  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Maxwel C Oliveira
    affiliation: '1'
    etal: Oliveira # First author's last name. 
  - name: Amit J Jhala
    affiliation: '2'
  - name: Mark Bernards
    affiliation: '3'
  - name: Chris Proctor
    affiliation: '2'
  - name: Strahinja Stepanovic
    affiliation: '2'
  - name: Rodrigo Werle
    affiliation: '1'
    email: rwerle@uwisc.edu # Indicates corresponding Author
    institution: Department of Agronomy, University of Wisconsin-Madison, United States
    street: 1575 Linden Dr
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    zip: 53705
    country: United States
affiliation:
  - id: '1'
    department: Department of Agronomy
    institution: University of Wisconsin-Madison
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    country: United States
  - id: '2' 
    department: Department of Agronomy and Horticulture
    institution: University of Nebraska-Lincoln
    city: Lincoln
    state: Nebraska # only USA, Australia, Canada
    country: United States
  - id: '3' 
    department: Department of Agronomy
    institution: Western Illinois University
    city: Macomb
    state: Illnois # only USA, Australia, Canada
    country: United States
output: rticles::frontiers_article
csl: frontiers.csl
bibliography: adaptation.bib
---

@@

%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\tiny
 \keyFont{ \section{Keywords:} Text Text Text Text Text Text Evolution Weed }

@@

## Subsection 1 {-}

@@

## Subsection 2 {-}

@@

MCO design, wrote,

<!--

When determining authorship the following criteria should be observed:

 - Substantial contributions to the conception or design of the work; or the
   acquisition, analysis, or interpretation of data for the work; AND
 - Drafting the work or revising it critically for important intellectual
   content; AND
 - Final approval of the version to be published ; AND
 - Agreement to be accountable for all aspects of the work in ensuring that
   questions related to the accuracy or integrity of any part of the work are
   appropriately investigated and resolved.

Contributors who meet fewer than all 4 of the above criteria for authorship
should not be listed as authors, but they should be acknowledged.
(http://www.icmje.org/roles_a.html)

-->

The statement about the authors and contributors can be up to several sentences
long, describing the tasks of individual authors referred to by their initials
and should be included at the end of the manuscript before the References
section.

@@

Funding:

@@

```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "150mm", out.height = "260mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

>>>>

---
title: "Palmer amaranth (*Amaranthus palmeri*) adaptation to agroecossystems"
runtitle: "Palmer amaranth adaptation"
documentclass: frontiersSCNS  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Maxwel C Oliveira
    affiliation: '1'
    etal: Oliveira # First author's last name. 
  - name: Amit J Jhala
    affiliation: '2'
  - name: Mark Bernards
    affiliation: '3'
  - name: Chris Proctor
    affiliation: '2'
  - name: Strahinja Stepanovic
    affiliation: '2'
  - name: Rodrigo Werle
    affiliation: '1'
    email: rwerle@uwisc.edu # Indicates corresponding Author
    institution: Department of Agronomy, University of Wisconsin-Madison, United States
    street: 1575 Linden Dr
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    zip: 53705
    country: United States
affiliation:
  - id: '1'
    department: Department of Agronomy
    institution: University of Wisconsin-Madison
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    country: United States
  - id: '2' 
    department: Department of Agronomy and Horticulture
    institution: University of Nebraska-Lincoln
    city: Lincoln
    state: Nebraska # only USA, Australia, Canada
    country: United States
  - id: '3' 
    department: Department of Agronomy
    institution: Western Illinois University
    city: Macomb
    state: Illnois # only USA, Australia, Canada
    country: United States
output: rticles::frontiers_article
csl: frontiers.csl
bibliography: adaptation.bib
---

@@

%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\tiny
 \keyFont{ \section{Keywords:} Evolution Flowering Management Pigweed Weed}

@@

## Cumulative flowering {-}

@@

## Palmer amaranth height and biomass {-}

@@

RW: designed the experiments; AJ, CP, MB, MO, and SS: conducted the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB, MO, SS, and RW: conceptualized the research. All authors reviewed the manuscript.

@@

Funding: This work received no specific grant from any funding agency, commercial, or not-for-profit sectors

@@

```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "150mm", out.height = "200mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

### 2021-07-26 Frontiers/Frontiers.tex
as

\title[Short Title]{Adaptation of Palmer amaranth (\emph{Amaranthus
palmeri}) to agroecossystems}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

@@

Light and temperature are likely the main environment requirements for
Palmer amaranth successful adaptation. Palmer amaranth is reported with
an extended germination period (Jha et al., 2010). Germination of Palmer
amaranth is triggered by 18 C soil temperature (Keeley et al., 1987),
and optimal germination and biomass production occur at 35/30 C day and
night temperatures (Guo and Al-Khatib, 2003). Water has not shown to
limit Palmer amaranth fitness. Under continuous water stress, Palmer
amaranth survived and produced at least 14000 seeds plant-1 (Chahal et
al., 2018). Also, seeds from Palmer amaranth growing under water stress
conditions were heavier, less dormant, and prompt for germination
(Matzrafi et al., 2021). The continuous global temperature warming can
impact agriculture and promote niches for Palmer amaranth
invasion/adaptation into new environments. Currently, it is estimated
that the greatest climatic risk of Palmer amaranth establishment is
agronomic crops in Australia and Sub-Sahara Africa (Kistner and
Hatfield, 2018). Temperature is a key factor limiting Palmer amaranth
expansion to cooler geographies (Briscoe Runquist et al., 2019);
however, under future climate change Palmer amaranth is likely to expand
northward into Canada and Northern Europe (Kistner and Hatfield, 2018;
Briscoe Runquist et al., 2019).

Palmer amaranth is already documented in agronomic crops of South
America (Larran et al., 2017; Küpper et al., 2017) and Southern Europe
(Milani et al., 2021). In the US, Palmer amaranth is established at crop
(Garetson et al., 2019) and non-crop land (Bagavathiannan and
Norsworthy, 2016) in the warm southern United States but its range is
expanding to cool temperatures northward. For example, herbicide
resistant Palmer amaranth is widespread in Nebraska (Oliveira et al.,
2021), Michigan (Kohrt et al., 2017), and Connecticut (Aulakh et al.,
2021). Successful cases of Palmer amaranth invasion and near to
eradication is reported in Minnesota (Yu et al., 2021). No Palmer
amaranth actively growing was found in Canada; however, Palmer amaranth
seeds was detected in sweet potato slips (Page et al., 2021).
Nonetheless, it seems fated to manage Palmer amaranth in agronomic crops
throughout multiple environments in the near future. Therefore,
strategies on Palmer amaranth management should encompass the
agroecosystem level but not attempts to eradicate the weed. Most tactics
to manage Palmer amaranth are based technology fixes (Scott, 2011),
which are short-term (e.g., herbicide and/or tillage) rather than
long-term ecological management.

@@

\hypertarget{subsection-1}{%
\subsection*{Subsection 1}\label{subsection-1}}
\addcontentsline{toc}{subsection}{Subsection 1}

@@

\hypertarget{subsection-2}{%
\subsection*{Subsection 2}\label{subsection-2}}
\addcontentsline{toc}{subsection}{Subsection 2}

@@

MCO design, wrote,

The statement about the authors and contributors can be up to several
sentences long, describing the tasks of individual authors referred to
by their initials and should be included at the end of the manuscript
before the References section.

@@

{\centering \includegraphics[width=150mm,height=260mm]{../data analysis/figures/combine 3}

>>>>

\title[Palmer amaranth adaptation]{Palmer amaranth (\emph{Amaranthus
palmeri}) adaptation to agroecossystems}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

@@

Light and temperature are likely the main environment requirements for
Palmer amaranth successful adaptation. Palmer amaranth is reported with
an extended germination period (Jha et al., 2010). Germination of Palmer
amaranth is triggered by 18 C soil temperature (Keeley et al., 1987),
and optimal germination and biomass production occur at 35/30 C day and
night temperatures (Guo and Al-Khatib, 2003). Water has not shown to
limit Palmer amaranth fitness. Under continuous water stress, Palmer
amaranth survived and produced at least 14000 seeds plant-1 (Chahal et
al., 2018). Also, seeds from Palmer amaranth growing under water stress
conditions were heavier, less dormant, and prompt for germination
(Matzrafi et al., 2021). The continuous global temperature warming can
impact agriculture and promote niches for Palmer amaranth
invasion/adaptation into new environments. Currently, it is estimated
that the greatest climatic risk of Palmer amaranth establishment are
agronomic crops in Australia and Sub-Sahara Africa (Kistner and
Hatfield, 2018). Temperature is a key factor limiting Palmer amaranth
expansion to cooler geographies (Briscoe Runquist et al., 2019);
however, under future climate change Palmer amaranth is likely to expand
northward into Canada and Northern Europe (Kistner and Hatfield, 2018;
Briscoe Runquist et al., 2019).

Palmer amaranth is already documented in agronomic crops of South
America (Larran et al., 2017; Küpper et al., 2017) and Southern Europe
(Milani et al., 2021). In the US, Palmer amaranth is established at crop
(Garetson et al., 2019) and non-crop land (Bagavathiannan and
Norsworthy, 2016) in the warm southern United States but its range is
expanding to cool temperatures northward. For example, herbicide
resistant Palmer amaranth is widespread in Nebraska (Oliveira et al.,
2021), Michigan (Kohrt et al., 2017), and Connecticut (Aulakh et al.,
2021). Successful cases of Palmer amaranth invasion and near to
eradication is reported in Minnesota (Yu et al., 2021). No Palmer
amaranth actively growing was found in Canada; however, Palmer amaranth
seeds was detected in sweet potato slips (Page et al., 2021).
Nonetheless, it seems fated the need to manage Palmer amaranth in
agronomic crops throughout multiple environments in the near future.
Therefore, strategies on Palmer amaranth management should encompass the
agroecosystem level but not only attempts to eradicate the weed. Most
tactics to manage Palmer amaranth are based on technology fixes (Scott,
2011), which are short-term (e.g., herbicide and/or tillage) rather than
long-term weed management.

@@

\hypertarget{cumulative-flowering}{%
\subsection*{Cumulative flowering}\label{cumulative-flowering}}
\addcontentsline{toc}{subsection}{Cumulative flowering}

@@

\hypertarget{palmer-amaranth-height-and-biomass}{%
\subsection*{Palmer amaranth height and
biomass}\label{palmer-amaranth-height-and-biomass}}
\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}

@@

RW: designed the experiments; AJ, CP, MB, MO, and SS: conducted the
experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB,
MO, SS, and RW: conceptualized the research. All authors reviewed the
manuscript.

@@

{\centering \includegraphics[width=150mm,height=200mm]{../data analysis/figures/combine 3}
